Exploration

### 2018-03-01 exploration.Rmd
moved functions to separate script; cleaned code

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pkgs <- c("fields", "sevcheck", "gbPopMod", "IPMpack", "tidyverse", "magrittr")
suppressMessages(invisible(lapply(pkgs, library, character.only=TRUE)))
theme_set(theme_bw())
lc.df <- read_csv("data/landcover_5km.csv") %>% 
  filter(!is.na(bio1_mean))
```

@@

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_` for survival, `g_` for growth, etc) and the type of covariate (`_z` for size, `_x` for environmental covariates).
```{r sp_i}
#set.seed(20)
subsamp <- NA
mean_abund <- 2000
regPlots <- F
ipmPlots <- F
if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
} else { cells <- sample(1:n_distinct(lc.df$CellID), subsamp, replace=FALSE) }
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-4, .75), # b1 + b2*z
                    s_x=c(1, -.1, 1, -.1), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(1, -.1, 1, -.1), #b1*x1 + ...
                    g_sig=1, # N(E, sig)
                    fl_z=c(-3, .6), # b1 + b2*z
                    fl_x=c(1, -.01, 1, -.01), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, 1, -.1), # b1*x1 + ...
                    rcrt=c(1.5, 0.4), #N(mean=rcrt1, sd=rcrt2)
                    p_est=0.02))
```

@@

## Simulation: IPM framework
```{r S}
########
## set up
########
# set up storage objects
p <- sp_i$p
d <- vector("list", length=n.cell)
lambdas <- lambdas2 <- rep(NA, n.cell)

@@

# life history functions
z_pow <- function(z.vec, n_z) {
  if(n_z==1) {
    return(matrix(1, ncol=1, nrow=length(z.vec)))
  } else {
    z <- matrix(z.vec, nrow=length(z.vec), ncol=n_z-1)
    for(i in 1:(n_z-1)) z[,i] <- z[,i]^i
    z <- cbind(1, z)
    return(z)
  }
}
calc_surv <- function(z.v, n_sz, p, X.s) {
  z <- z_pow(z.v, n_sz)
  u <- exp(z %*% p$s_z + c(t(X.s) %*% p$s_x))
  return(u / (1+u))
}
calc_grow <- function(z1, z.v, n_gz, p, X.g) {
  z <- z_pow(z.v, n_gz)
  g <- dnorm(z1, mean=z %*% p$g_z + c(t(X.g) %*% p$g_x), 
             sd=p$g_sig)
  return(g)
}
calc_flwr <- function(z.v, n_flz, p, X.fl) {
  z <- z_pow(z.v, n_flz)
  u <- exp(z %*% p$fl_z + c(t(X.fl) %*% p$fl_x))
  return(u / (1+u))
}
calc_seed <- function(z1, z.v, n_seedz, n_flz, p, X.seed, X.fl) {
  z <- z_pow(z.v, n_seedz)
  calc_flwr(z.v, n_flz, p, X.fl) * p$p_est *
      dnorm(z1, p$rcrt[1], p$rcrt[2]) *
      exp(z %*% p$seed_z + c(t(X.seed) %*% p$seed_x))
}

########
## loop through cells
########
for(i in 1:n.cell) {
  d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=12))
  lo <- min(d[[i]]$size)*0.8
  hi <- max(d[[i]]$size)*1.1
  size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)

  # generate data
  ## survival
  E_surv <- antilogit(size.mx[,1:n_sz] %*% p$s_z + c(X_s[i,] %*% p$s_x))
  d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)

  ## growth
  E_grow <- size.mx[,1:n_gz] %*% p$g_z + c(X_g[i,] %*% p$g_x)
  d[[i]]$sizeNext <- rnorm(n_i[i], E_grow, p$g_sig)
  d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)

  ## flowering 
  E_fl <- antilogit(size.mx[,1:n_flz] %*% p$fl_z + c(X_fl[i,] %*% p$fl_x))
  d[[i]]$fl <- rbinom(n_i[i], 1, E_fl)

  ## seed production
  E_seed <- d[[i]]$fl * exp((size.mx[,1:n_seedz] %*% p$seed_z + 
                               c(X_seed[i,] %*% p$seed_x)))
  d[[i]]$seed <- rpois(n_i[i], E_seed)
  d[[i]]$seed <- ifelse(d[[i]]$fl, d[[i]]$seed, NA)
  seed.range <- range(d[[i]]$seed, na.rm=TRUE)

  ## recruits
  rcrt.n <- max(sum(d[[i]]$seed, na.rm=TRUE)*p$p_est, 2)
  rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
  d[[i]][rcrt.i,] <- NA
  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcrt[1], p$rcrt[2])

  # fill IPM matrices
  ## Manual: use assigned slopes
  n <- 50  # ncells in matrix
  b <- lo + c(0:n)*(hi - lo)/n  # boundary points
  y <- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
  h <- y[2] - y[1]  # step size

  G.mx <- h*outer(y, y, calc_grow, n_gz=n_gz, p=p, X.g=X_g[i,])
  S.v <- calc_surv(y, n_sz, p, X_s[i,])
  F.mx <- h*outer(y, y, calc_seed, n_seedz=n_seedz, p=p, n_flz=n_flz,
                  X.seed=X_seed[i,], X.fl=X_fl[i,])
  P.mx <- G.mx
  for(k in 1:(n/2)) { # Fix eviction of individuals below/above size limits
    G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
    P.mx[,k] <- G.mx[,k]*S.v[k]
  }
  for(k in (n/2+1):n) {
    G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
    P.mx[,k] <- G.mx[,k]*S.v[k]
  }
  IPM <- P.mx + F.mx

  ## IPMpack: estimate slopes from d[[i]]
  so <- makeSurvObj(d[[i]], surv ~ size)
  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
  fo <- makeFecObj(d[[i]],
                   Formula=c(fl~size, seed~size+size2),
                   Family=c("binomial", "poisson"),
                   Transform=c("none", "none"),
                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est))
  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, minSize=lo, maxSize=hi,
                         correction="constant")
  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
                          correction="constant")
  IPM2 <- P.mx2 + F.mx2

  lambdas[i] <- Re(eigen(IPM)$values[1])
  lambdas2[i] <- Re(eigen(IPM2)$values[1])

  # Plots
  z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
  if(regPlots) {
    par(mfrow=c(3,2), mar=c(4,4,2,1))
    xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      points(z.t$size, E_surv, cex=0.2, col="red")
      lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      points(z.t$size, E_grow, cex=0.2, col="red")
      lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      points(z.t$size, E_fl, cex=0.2, col="red")
      lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      points(z.t$size, E_seed, cex=0.2, col="red")
      lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
    w1 <- Re(eigen(IPM)$vectors[,1]); stable.dist1 <- w1/sum(w1)
    w2 <- Re(eigen(IPM2)$vectors[,1]); stable.dist2 <- w2/sum(w2)
    yl <- range(c(stable.dist1, stable.dist2))
    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    lines(y, stable.dist2, lwd=2, lty=2, col="blue")
    # plot(y, S.v); points(y, apply(P.mx, 2, sum), col="red", type="l")
    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
      xxx <- seq(0, hi*.5, length.out=100)
      lines(xxx, dnorm(xxx, p$rcrt[1], p$rcrt[2]), col="red")
  }
  if(ipmPlots) {
    par(mfrow=c(2,2), mar=c(4,4,2,1)); main.i <- paste("Cell", i)
    # sens.mx <- sens(IPM2); elas.mx <- elas(IPM2)
    sens.mx <- sens(IPM); elas.mx <- elas(IPM)
    image.plot(y, y, t(P.mx), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
      abline(a=0, b=1, col="white", lty=2)
    image.plot(y, y, t(P.mx2), main=main.i, xlab="Size (t)", ylab="Size (t+1)")
      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2))
      abline(a=0, b=1, col="white", lty=2)
    image.plot(1:dim(sens.mx)[1], 1:dim(sens.mx)[1], t(sens.mx),
               main=paste("Sens:", main.i), xlab="Size (t)", ylab="Size (t+1)")
    image.plot(1:dim(elas.mx)[1], 1:dim(elas.mx)[1], t(elas.mx),
               main=paste("Elas:", main.i), xlab="Size (t)", ylab="Size (t+1)")
  }
  if(i %% 100 == 0) cat("Finished cell", i, "\n")
}
lam.df <- env.df %>% add_column(lambda=lambdas, lambda2=lambdas2)
ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
ggplot(lam.df, aes(x=x, y=y, fill=log(lambdas2))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
par(mfrow=c(2,2))
plot(log(lambdas) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambdas) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambdas), log(lambdas2)); abline(v=0, h=0, a=0, b=1, lty=c(3,2,2))
plot(density(log(lambdas))); abline(v=0, col="red", lty=2)
```

@@

```
### Mechanistic: IPM
```{r P_IPM}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pkgs <- c("fields", "sevcheck", "gbPopMod", "IPMpack", "tidyverse", "magrittr")
suppressMessages(invisible(lapply(pkgs, library, character.only=TRUE)))
theme_set(theme_bw())
source("code/fn_IPM.R"); source("code/fn_aux.R")
lc.df <- read_csv("data/landcover_5km.csv") %>% 
  filter(!is.na(bio1_mean))
```

@@

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
subsamp <- NA
mean_abund <- 200
regPlots <- F
ipmPlots <- F
n <- 100  # ncells in IPM matrix
if(is.na(subsamp)) { cells <- 1:n_distinct(lc.df$CellID)
} else { cells <- sample(1:n_distinct(lc.df$CellID), subsamp, replace=FALSE) }
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
                    s_x=c(1, -.1, -.5, -.1, .05), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(1, -.1, 1.5, -.1), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
                    fl_x=c(1, -.01, -1, -.01, .05), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, 1, -.1, .05), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.02, # p(establishment)
                    rcr_SB=0.2, # p(recruit from seedbank)
                    rcr_dir=0.6, # p(recruit directly)
                    s_SB=0.3 # p(survive in seedbank additional year)
                    ))
```

@@

## Simulation: IPM framework
```{r S}
########
## set up
########
# set up storage objects
p <- sp_i$p
IPMs <- array(0, dim=c(n+1, n+1, n.cell))
d <- vector("list", length=n.cell)
lambdas <- B_1 <- rep(NA, n.cell)
B <- rpois(n.cell, n_i)

@@

########
## loop through cells
########
for(i in 1:n.cell) {
  d[[i]] <- data.frame(size=runif(n=n_i[i], min=1, max=12))
  lo <- min(d[[i]]$size)*0.75
  hi <- max(d[[i]]$size)*1.25
  size.mx <- cbind(1, d[[i]]$size, d[[i]]$size^2, d[[i]]$size^3)

  #---- generate data
  ## survival
  E_surv <- antilogit(size.mx[,1:n_sz] %*% p$s_z + c(X_s[i,] %*% p$s_x))
  d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)
  ## growth
  E_grow <- size.mx[,1:n_gz] %*% p$g_z + c(X_g[i,] %*% p$g_x)
  d[[i]]$sizeNext <- rnorm(n_i[i], E_grow, p$g_sig)
  d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)
  ## flowering 
  E_fl <- antilogit(size.mx[,1:n_flz] %*% p$fl_z + c(X_fl[i,] %*% p$fl_x))
  d[[i]]$fl <- rbinom(n_i[i], 1, E_fl)
  ## seed production
  E_seed <- d[[i]]$fl * 
    exp((size.mx[,1:n_seedz] %*% p$seed_z + c(X_seed[i,] %*% p$seed_x)))
  d[[i]]$seed <- rpois(n_i[i], E_seed)
  d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
  d[[i]]$seed <- ifelse(d[[i]]$fl==1, d[[i]]$seed, NA)
  ## seed bank
  B_1[i] <- B[i] * p$s_SB * (1 - p$rcr_SB) + 
    sum(d[[i]]$seed, na.rm=TRUE) * (1 - p$rcr_dir) * p$s_SB
  ## recruits
  rcrt.n <- max((sum(d[[i]]$seed, na.rm=TRUE) * p$rcr_dir * p$p_est +
                   B[i] * p$rcr_SB * p$p_est), 
                2) # need ≥ 2 for IPMpack estimate of rcr_z'_mn, rcr_z'_sd
  rcrt.i <- (n_i[i]+1):(n_i[i]+rcrt.n)
  d[[i]][rcrt.i,] <- NA
  d[[i]]$sizeNext[rcrt.i] <- rnorm(rcrt.n, p$rcr_z[1], p$rcr_z[2])

  #---- fill IPM matrices
  ## Underlying Truth: use assigned slopes
  b <- lo + c(0:n)*(hi - lo)/n  # boundary points
  y <- 0.5*(b[1:n] + b[2:(n+1)])  # mesh points
  h <- y[2] - y[1]  # step size
  z.i <- 2:(n+1)  # continuous stage matrix indices
  ### growth and survival
  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_gz, X.g=X_g[i,])
  S.v <- calc_surv(y, p=p, n_sz=n_sz, X.s=X_s[i,])
  ### Fix eviction of individuals below/above size limits
  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
  ### fill P, F, and K
  P.mx <- F.mx <- matrix(0, ncol=n+1, nrow=n+1)
  for(k in z.i) P.mx[k,z.i] <- G.mx[k-1,]*S.v
  P.mx[1,1] <- calc_staySB(p)
  F.mx[z.i,z.i] <- h*outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                           X.seed=X_seed[i,], X.fl=X_fl[i,])
  F.mx[1,z.i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                            X.seed=X_seed[i,], X.fl=X_fl[i,])
  F.mx[z.i,1] <- calc_rcrSB(y, p)
  IPMs[,,i] <- P.mx + F.mx

  #---- progress
  if(i %% 100 == 0) cat("Finished cell", i, "\n")
}
lam.df <- env.df
lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
ggplot(lam.df, aes(x=x, y=y, fill=log(lambda))) + geom_tile() +
  scale_fill_gradient2(low="darkblue", mid="white", high="darkred", midpoint=0)
par(mfrow=c(2,2))
plot(log(lambda) ~ temp, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ prec, data=lam.df); abline(h=0, col="red", lty=2)
plot(log(lambda) ~ pOpn, data=lam.df); abline(h=0, col="red", lty=2)
plot(density(log(lam.df$lambda))); abline(v=0, col="red", lty=2)

# Plots
if(regPlots) {
  par(mfrow=c(3,2), mar=c(4,4,2,1))
  xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
  xxx <- seq(0, hi*.5, length.out=100)
  for(i in 1:n.cell) {
    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      points(z.t$size, E_surv, cex=0.2, col="red")
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      points(z.t$size, E_grow, cex=0.2, col="red"); abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      points(z.t$size, E_fl, cex=0.2, col="red")
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      points(z.t$size, E_seed, cex=0.2, col="red")
    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", 
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
  }
}
if(ipmPlots) {
  par(mfrow=c(2,2), mar=c(4,4,2,1))
  x <- c(0, y)
  for(i in 1:n.cell) {
    z.t <- d[[i]][!is.na(d[[i]]$size),]; main.i <- paste("Cell", i)
    z.t$y <- y[findInterval(z.t$size, y, all.inside=TRUE)]
    totSeeds <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=TRUE))
    image.plot(x, x, t(P.mx), main=paste("P matrix:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
      points(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,1/(n_i[i]^.25)))
      abline(a=0, b=1, col="white", lty=2)
    image.plot(x, x, t(F.mx), main=paste("F matrix:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
      points(totSeeds$y, totSeeds$seed/max(totSeeds$seed)*hi*.9)
    image.plot(x, x, t(sens(IPMs[,,i])), main=paste("Sensitivity:", main.i), 
               xlab="Size (t)", ylab="Size (t+1)")
    image.plot(x, x, t(elas(IPMs[,,i])), main=paste("Elasticity:", main.i),
               xlab="Size (t)", ylab="Size (t+1)")
  }
}
```

@@

```
### Mechanistic: IPM
```{r P_IPM}
for(i in 1:ncell) {
  # ## IPMpack: estimate slopes from d[[i]]
  # ### set up seedbank transitions
  d[[i]]$stage <- d[[i]]$stageNext <- "continuous"
  d[[i]]$stage[is.na(d[[i]]$size)] <- NA
  d[[i]]$stageNext[d[[i]]$surv==0] <- "dead"
  d[[i]]$number <- 1
  seedbank <- data.frame(size=NA, surv=1, sizeNext=NA,
                         fl=NA, seed=NA,
                         stage=c("seedbank", "seedbank", "continuous"),
                         stageNext=c("seedbank", "continuous", "seedbank"),
                         number=c(p$s_SB * (1 - p$rcr_SB), 
                                  p$rcr_SB * p$p_est, 
                                  (1 - p$rcr_dir) * p$s_SB))
  d[[i]] <- rbind(d[[i]], seedbank)
  d[[i]]$stage <- as.factor(d[[i]]$stage)
  d[[i]]$stageNext <- as.factor(d[[i]]$stageNext)
  ### construct kernels
  so <- makeSurvObj(d[[i]], surv ~ size)
  go <- makeGrowthObj(d[[i]], sizeNext ~ size + size2)
  fo <- makeFecObj(d[[i]],
                   Formula=c(fl~size+size2, seed~size+size2),
                   Family=c("binomial", "poisson"),
                   Transform=c("none", "none"),
                   meanOffspringSize=mean(
                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
                                       !is.na(d[[i]]$sizeNext)]),
                   sdOffspringSize=sd(
                     d[[i]]$sizeNext[is.na(d[[i]]$size) &
                                       !is.na(d[[i]]$sizeNext)]),
                   fecConstants=data.frame(seedlingEstablishmentRate=p$p_est),
                   offspringSplitter=data.frame(seedbank=(1-p$rcr_dir),
                                                continuous=p$rcr_dir)
                   )
  F.mx2 <- makeIPMFmatrix(fecObj=fo, minSize=lo, maxSize=hi,
                          correction="constant")
  D.mx2 <- makeDiscreteTrans(d[[i]], 
                             continuousToDiscreteExplanatoryVariables=1)
  dummy <- as.matrix(fo@offspringRel$coefficients[1], list(1, "seedbank"))
  D.mx2@meanToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
  dummy <- as.matrix(fo@sdOffspringSize, dimnames=list(1, "seedbank"))
  D.mx2@sdToCont <- as.matrix(dummy, dimnames=c(1, "seedbank"))
  D.mx2@discreteTrans[1,1] <- p$s_SB * (1 - p$rcr_SB)
  D.mx2@discreteTrans[3,1] <- 1 - sum(D.mx2@discreteTrans[1:2,1])
  P.mx2 <- makeIPMPmatrix(growObj=go, survObj=so, discreteTrans=D.mx2,
                            minSize=lo, maxSize=hi, correction="constant")
  IPM2 <- P.mx2 + F.mx2
  lambdas2[i] <- Re(eigen(IPM2)$values[1])

  if(regPlots) {
    par(mfrow=c(3,2), mar=c(4,4,2,1))
    xx <- tibble(size=seq(lo, hi, length.out=100), size2=size^2, size3=size^3)
    plot(surv~size, data=d[[i]], ylim=c(0,1), pch=20, col=rgb(0,0,0,0.2),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      points(z.t$size, E_surv, cex=0.2, col="red")
      # lines(xx$size, predict(so@fit, xx, type="response"), col="blue")
    plot(sizeNext~size, data=z.t, pch=20, col=rgb(0,0,0,0.2), 
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      points(z.t$size, E_grow, cex=0.2, col="red")
      # lines(xx$size, predict(go@fit, xx, type="response"), col="blue")
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=rgb(0,0,0,0.2),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      points(z.t$size, E_fl, cex=0.2, col="red")
      # lines(xx$size, predict(fo@fitFec[[1]], xx, type="response"), col="blue")
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=rgb(0,0,0,0.2), 
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      points(z.t$size, E_seed, cex=0.2, col="red")
      # lines(xx$size, predict(fo@fitFec[[2]], xx, type="response"), col="blue")
    w1 <- Re(eigen(IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    # w2 <- Re(eigen(IPM2)$vectors[,1])[-1]; stable.dist2 <- w2/sum(w2)
    yl <- range(stable.dist1)#range(c(stable.dist1, stable.dist2))
    plot(y, stable.dist1, ylab="", type="l", lwd=2, col="red", ylim=yl,
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    # lines(y, stable.dist2, col="blue")
    hist(d[[i]]$sizeNext[rcrt.i], main=paste("Recruit sizes:", main.i), 
         xlab="Recruit size (t+1)", col="grey", freq=FALSE)
      xxx <- seq(0, hi*.5, length.out=100)
      rcrt.z <- d[[i]]$sizeNext[is.na(d[[i]]$size) & !is.na(d[[i]]$sizeNext)]
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red", lwd=2)
      # lines(xxx, dnorm(xxx, mean(rcrt.z), sd(rcrt.z)), col="blue")
  }
}
